Decoupled Fallback: A Portable Single-Pass GPU Scan

### 2025-02-26 paper.tex
making a couple of tables fit to page/column width

\section{Evaluation}
\subsection{Experimental Setup}
We configure workgroups with a size $w = 256$ threads, where each thread processes $p = 4$ vectors of size 4, resulting in 16 elements per thread and a total work tile size of $t = 4096$. We find a maximum spin count of 4 effectively balances responsiveness to blocking while minimizing false-positive fallbacks. As the size of the local spine is inversely proportional to the subgroup size, we allocate sufficient shared memory to accommodate the WebGPU standard's minimum subgroup size of 4. Due to our overall thriftiness with shared memory, total usage remains modest at only $\sim$1 KB\@. For all tests, we use an input size of $2^{25}$ and an inclusive prefix sum. Our test hardware, listed POSITIONAL-RELATION, spans a wide range of FPG and performance capabilities. We empirically find that Google Chrome's native implementation of WebGPU, Dawn, provides the best performance and use it for all testing, except on the Arm-Mali GPU, which requires Android native, necessitating the use of WGPU\@.\footnote{To keep the perforamance as close to Dawn as possible, we translate our WGSL code with Dawn's Tint transpiler, and provide the raw SPIR-V to WGPU\@.}
\begin{table}
  \centering
  \small
  \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} l l c r r}
    \toprule
    Vendor & Device Name & LOBE & Occupancy & Subgroup Size \\
    \midrule
    Intel  & HD Graphics 620        & \textbf{Yes} & 9   & 8--32  \\
    ARM    & Mali-G78 MP20          & \textbf{No}  & 40  & 16  \\
    Apple  & M3     (10 Cores)      & \textbf{No}  & 80  & 32  \\
    Apple  & M1 Max (32 Cores)      & \textbf{No}  & 160 & 32  \\
    NVIDIA & GeForce RTX 2080 Super & \textbf{Yes} & 192 & 32  \\
    AMD    & Radeon RX 7900 XT      & \textbf{Yes} & 420 & 32--64  \\
    \bottomrule
  \end{tabular*}
  \caption{Hardware Specifications of Testing Devices.}
  \label{tab:hardware}
\end{table}

@@

\begin{table*}
  \centering
  \begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} l r r r r r r}
    \toprule
    Metric & Intel HD620 & ARM Mali-G78 MP20 & Apple M3 & Apple M1 Max & Nvidia 2080 Super & AMD 7900 XT \\
    \midrule
    Avg \emph{Memcpy} (32-bit ele/sec $\cdot 10^9$)  & 1.50   & 2.48  & 10.25  & 38.62  & 54.39  & 72.28  \\
    Avg \emph{DF} (32-bit ele/sec $\cdot 10^9$)   & 1.49   & 2.79  & 10.22  & 38.84  & 53.29  & 94.60  \\
    Avg \emph{RTS} (32-bit ele/sec $\cdot 10^9$)      & 1.06   & 2.02  & 6.81   & 25.13  & 35.90  & 61.49  \\
    \emph{DF}/\emph{Memcpy}       & 99.1\%    & 112.8\%  & 99.8\%    & 100.6\%   & 98.0\%    & 130.9\%   \\
    \emph{DF}/\emph{RTS}          & 140.2\%   & 138.5\%  & 150.2\%   & 154.5\%   & 148.4\%   & 153.8\%   \\
    Avg Spins per Wg    & 0.305    & 1.477    & 0.606    & 0.604    & 0.419    & 0.148    \\
    Avg Lookback Length per Wg  & 1.726    & 3.537    & 3.415    & 10.782   & 13.046   & 20.722   \\
    Avg Total Fallbacks Initiated      & 254.662   & 2504.704  & 365.546   & 990.962   & 86.400   & 9.308   \\
    Avg Total Successful Insertions    & 6.170     & 122.436   & 4.612     & 0.012    & 0.284    & 0.020    \\
    \bottomrule
  \end{tabular*}
  \caption{\emph{Decoupled Fallback} performance, 32-bit Inclusive Prefix Sum, $\mathbf{2^{25}}$ 32-bit elements}
  \label{tab:results}
\end{table*}

@@

\subsection{Baseline Comparisons: \emph{Memcpy}}
\label{sec:memcpy}
\begin{table}
  \small
  \centering
  \begin{tabular}{l r r}
    \toprule
    Device                                & \makecell{Memory                   \\ Bandwidth \\ (GB/s)} & \makecell{Expected \\ Throughput \\ (32-bit ele/sec)} \\
    \midrule
    Intel HD620 Single-Channel            & 17.1             & 2.13e9          \\
    Intel HD620 Dual-Channel              & 34.1             & 4.26e9          \\
    ARM Mali-G78 MP20 / Tensor (Pixel 6a) & 51.2             & 6.40e9          \\
    Apple M3 (??? Cores)                  & 102.4            & 12.8e9          \\
    Apple M1 Max (32 Cores)               & 409.6            & 51.2e9          \\
    Nvidia 2080 Super                     & 496              & 62.0e9          \\
    AMD 7900 XT                           & 800              & 100.0e9         \\
    \bottomrule
  \end{tabular}
  \caption{Vendor-Specified Memory Bandwidth and Expected Throughput (32-bit Elements per Second).}
  \label{tab:memory_bandwidth}
\end{table}
Although all of Dawn's backend APIs support \emph{timestamp queries inside encoders}, Apple GPUs do not, so we are forced to implement our own \emph{Memcpy} kernel to measure memory bandwidth, as opposed to simply timing an API's copy buffer operation. It is well known that \emph{observed} peak memory bandwidth, tends to diverge from vendor-listed peak bandwidth, so for completeness we list these figures above. We make a number of observations:
\begin{itemize}
  \item \textbf{Intel HD620:} As our HD620 laptop is only equipped with a single memory DIMM, its peak bandwidth is limited to the lower bound, 2.13e9 ele/sec. Thus, our observed bandwidth of 1.498e9 is reasonable.
  \item \textbf{ARM Mali-G78 MP20 / Pixel 6a (Tensor):} The observed throughput is significantly lower than expected, despite a vendor-listed bandwidth of 6.40e9 ele/sec. The cause of this discrepancy remains unclear, but we note other developers have observed similar speeds.
  \item \textbf{Apple M1 Pro Max:} Observed throughput matches expectations.
  \item \textbf{Nvidia 2080 Super:} Observed throughput matches expectations.
  \item \textbf{AMD 7900 XT:} Our above result, where \emph{DF} was observed to be faster than \emph{Memcopy}, is likely due to compute latency counterintuitively improving memory access patterns, leading to reduced paging. Furthermore, we see that our \emph{DF} throughput is within spec and reasonable.
\end{itemize}

>>>>

\section{Evaluation}
\subsection{Experimental Setup}
We configure workgroups with a size $w = 256$ threads, where each thread processes $p = 4$ vectors of size 4, resulting in 16 elements per thread and a total work tile size of $t = 4096$. We find a maximum spin count of 4 effectively balances responsiveness to blocking while minimizing false-positive fallbacks. As the size of the local spine is inversely proportional to the subgroup size, we allocate sufficient shared memory to accommodate the WebGPU standard's minimum subgroup size of 4. Due to our overall thriftiness with shared memory, total usage remains modest at only $\sim$1 KB\@. For all tests, we use an input size of $2^{25}$ and an inclusive prefix sum. Our test hardware, listed POSITIONAL-RELATION, spans a wide range of FPG and performance capabilities. We empirically find that Google Chrome's native implementation of WebGPU, Dawn, provides the best performance and use it for all testing, except on the Arm-Mali GPU, which requires Android native, necessitating the use of WGPU\@.\footnote{To keep the perforamance as close to Dawn as possible, we translate our WGSL code with Dawn's Tint transpiler, and provide the raw SPIR-V to WGPU\@.}
\begin{table}
  \centering
  \small
  \setlength{\tabcolsep}{1pt}
  \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} l l c r r}
    \toprule
    Vendor & Device Name & LOBE & Occupancy & Subgroup Size \\
    \midrule
    Intel  & HD Graphics 620        & \textbf{Yes} & 9   & 8--32  \\
    ARM    & Mali-G78 MP20          & \textbf{No}  & 40  & 16  \\
    Apple  & M3     (10 Cores)      & \textbf{No}  & 80  & 32  \\
    Apple  & M1 Max (32 Cores)      & \textbf{No}  & 160 & 32  \\
    NVIDIA & GeForce RTX 2080 Super & \textbf{Yes} & 192 & 32  \\
    AMD    & Radeon RX 7900 XT      & \textbf{Yes} & 420 & 32--64  \\
    \bottomrule
  \end{tabular*}
  \caption{Hardware Specifications of Testing Devices.}
  \label{tab:hardware}
\end{table}

@@

\begin{table*}
  \centering
  \setlength{\tabcolsep}{1pt}
  \begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} l r r r r r r}
    \toprule
    Metric & Intel HD620 & ARM Mali-G78 MP20 & Apple M3 & Apple M1 Max & Nvidia 2080 Super & AMD 7900 XT \\
    \midrule
    Avg \emph{Memcpy} (32-bit ele/sec $\cdot 10^9$)  & 1.50   & 2.48  & 10.25  & 38.62  & 54.39  & 72.28  \\
    Avg \emph{DF} (32-bit ele/sec $\cdot 10^9$)   & 1.49   & 2.79  & 10.22  & 38.84  & 53.29  & 94.60  \\
    Avg \emph{RTS} (32-bit ele/sec $\cdot 10^9$)      & 1.06   & 2.02  & 6.81   & 25.13  & 35.90  & 61.49  \\
    \emph{DF}/\emph{Memcpy}       & 99.1\%    & 112.8\%  & 99.8\%    & 100.6\%   & 98.0\%    & 130.9\%   \\
    \emph{DF}/\emph{RTS}          & 140.2\%   & 138.5\%  & 150.2\%   & 154.5\%   & 148.4\%   & 153.8\%   \\
    Avg Spins per Wg    & 0.305    & 1.477    & 0.606    & 0.604    & 0.419    & 0.148    \\
    Avg Lookback Length per Wg  & 1.726    & 3.537    & 3.415    & 10.782   & 13.046   & 20.722   \\
    Avg Total Fallbacks Initiated      & 254.662   & 2504.704  & 365.546   & 990.962   & 86.400   & 9.308   \\
    Avg Total Successful Insertions    & 6.170     & 122.436   & 4.612     & 0.012    & 0.284    & 0.020    \\
    \bottomrule
  \end{tabular*}
  \caption{\emph{Decoupled Fallback} performance, 32-bit Inclusive Prefix Sum, $\mathbf{2^{25}}$ 32-bit elements}
  \label{tab:results}
\end{table*}

@@

\subsection{Baseline Comparisons: \emph{Memcpy}}
\label{sec:memcpy}
\begin{table}
  \small
  \centering
  \begin{tabular}{l r r}
    \toprule
    Device                                & \makecell{Memory                   \\ Bandwidth \\ (GB/s)} & \makecell{Expected \\ Throughput \\ (32-bit ele/sec)} \\
    \midrule
    Intel HD620 Single-Channel            & 17.1             & 2.13e9          \\
    Intel HD620 Dual-Channel              & 34.1             & 4.26e9          \\
    ARM Mali-G78 MP20 / Tensor (Pixel 6a) & 51.2             & 6.40e9          \\
    Apple M3 (10 Cores)                   & 102.4            & 12.8e9          \\
    Apple M1 Max (32 Cores)               & 409.6            & 51.2e9          \\
    Nvidia 2080 Super                     & 496              & 62.0e9          \\
    AMD 7900 XT                           & 800              & 100.0e9         \\
    \bottomrule
  \end{tabular}
  \caption{Vendor-Specified Memory Bandwidth and Expected Throughput (32-bit Elements per Second).}
  \label{tab:memory_bandwidth}
\end{table}
Although all of Dawn's backend APIs support \emph{timestamp queries inside encoders}, Apple GPUs do not, so we are forced to implement our own \emph{Memcpy} kernel to measure memory bandwidth, as opposed to simply timing an API's copy buffer operation. It is well known that \emph{observed} peak memory bandwidth, tends to diverge from vendor-listed peak bandwidth, so for completeness we list these figures above. We make a number of observations:
\begin{itemize}
  \item \textbf{Intel HD620:} As our HD620 laptop is only equipped with a single memory DIMM, its peak bandwidth is limited to the lower bound, 2.13e9 ele/sec. Thus, our observed bandwidth of 1.498e9 is reasonable.
  \item \textbf{ARM Mali-G78 MP20 / Pixel 6a (Tensor):} The observed throughput is significantly lower than expected, despite a vendor-listed bandwidth of 6.40e9 ele/sec. The cause of this discrepancy remains unclear, but we note other developers have observed similar speeds.
  \item \textbf{Apple M1 Pro Max:} Observed throughput matches expectations.
  \item \textbf{Nvidia 2080 Super:} Observed throughput matches expectations.
  \item \textbf{AMD 7900 XT:} Our above result, where \emph{DF} was observed to be faster than \emph{Memcopy}, is likely due to compute latency counterintuitively improving memory access patterns, leading to reduced paging. Furthermore, we see that our \emph{DF} throughput is within spec and reasonable.
\end{itemize}
